Examining FDR correction issue in clusterProfiler - RNA-seq example

### 2024-09-09 analysis/analyse_rnaseq1.Rmd
adding error bars

pdf("fig1_ORAjac.pdf",width=7, height=7)
par(mar=c(c(5.1, 5.1, 2.1, 2.1) ))
plot(cov,jac2,xlim=c(0,30000),ylim=c(0,1),pch=19,col="lightgray",cex=3,
  xlab="gene coverage",ylab="Jaccard index")
text(cov,jac2,labels=names(cov))
mylm <- lm(jac2~cov)
abline(mylm,lty=2,lwd=2)
SLOPE=unname(signif(mylm$coefficients[2],3))
INT=unname(signif(mylm$coefficients[1],3))
PVAL=signif(summary(mylm)$coefficients[4],3)
RSQ=signif(summary(mylm)$r.squared,3)
HEADER=paste("Rsq =",RSQ,"; p-val =",PVAL)
mtext(HEADER)
dev.off()
```

@@

par(mar = c(5.1, 8.1, 4.1, 2.1))
library(beeswarm)
boxplot(unsl,horizontal=TRUE,las=1,col=c("lightblue","pink"),cex=0,xlab="no. significant sets FDR<0.05")
beeswarm(unsl,horizontal=TRUE,add=TRUE,col=c("blue","red"),pch=19,cex=0.6)

>>>>

pdf("fig1_ORAjac.pdf",width=7, height=7)
par(mar=c(c(5.1, 5.1, 2.1, 2.1) ))
plot(cov,jac2,xlim=c(0,30000),ylim=c(0,1),pch=19,col="lightgray",cex=3,
  xlab="gene coverage",ylab="Jaccard index")
text(cov,jac2,labels=names(cov))
mylm <- lm(jac2~cov)
abline(mylm,lty=2,lwd=2)
SLOPE=unname(signif(mylm$coefficients[2],3))
INT=unname(signif(mylm$coefficients[1],3))
PVAL=signif(summary(mylm)$coefficients[4],3)
RSQ=signif(summary(mylm)$r.squared,3)
HEADER=paste("Rsq =",RSQ,"; p-val =",PVAL)
mtext(HEADER)
dev.off()

# with SD
mysd <- apply(jac,2,function(x) { sd(x,na.rm=TRUE) } )
par(mar=c(c(5.1, 5.1, 2.1, 2.1) ))
plot(cov,jac2,xlim=c(0,30000),ylim=c(0,1),pch=19,col="lightgray",cex=3,
  xlab="gene coverage",ylab="Jaccard index")
text(cov,jac2,labels=names(cov))
mylm <- lm(jac2~cov)
abline(mylm,lty=2,lwd=2)
SLOPE=unname(signif(mylm$coefficients[2],3))
INT=unname(signif(mylm$coefficients[1],3))
PVAL=signif(summary(mylm)$coefficients[4],3)
RSQ=signif(summary(mylm)$r.squared,3)
HEADER=paste("Rsq =",RSQ,"; p-val =",PVAL)
mtext(HEADER)
arrows(x0=cov,x1=cov,y0=jac2+mysd,y1=jac2-mysd,code=3, angle=90, length=0.1 , col="gray40")

png("fig1_ORAjacsd.png")
par(mar=c(c(5.1, 5.1, 2.1, 2.1) ))
plot(cov,jac2,xlim=c(0,30000),ylim=c(0,1),pch=19,col="lightgray",cex=3,
  xlab="gene coverage",ylab="Jaccard index")
text(cov,jac2,labels=names(cov))
mylm <- lm(jac2~cov)
abline(mylm,lty=2,lwd=2)
SLOPE=unname(signif(mylm$coefficients[2],3))
INT=unname(signif(mylm$coefficients[1],3))
PVAL=signif(summary(mylm)$coefficients[4],3)
RSQ=signif(summary(mylm)$r.squared,3)
HEADER=paste("Rsq =",RSQ,"; p-val =",PVAL)
mtext(HEADER)
arrows(x0=cov,x1=cov,y0=jac2+mysd,y1=jac2-mysd,code=3, angle=90, length=0.1 , col="gray40")
dev.off()

pdf("fig1_ORAjacsd.pdf",width=7, height=7)
par(mar=c(c(5.1, 5.1, 2.1, 2.1) ))
plot(cov,jac2,xlim=c(0,30000),ylim=c(0,1),pch=19,col="lightgray",cex=3,
  xlab="gene coverage",ylab="Jaccard index")
text(cov,jac2,labels=names(cov))
mylm <- lm(jac2~cov)
abline(mylm,lty=2,lwd=2)
SLOPE=unname(signif(mylm$coefficients[2],3))
INT=unname(signif(mylm$coefficients[1],3))
PVAL=signif(summary(mylm)$coefficients[4],3)
RSQ=signif(summary(mylm)$r.squared,3)
HEADER=paste("Rsq =",RSQ,"; p-val =",PVAL)
mtext(HEADER)
arrows(x0=cov,x1=cov,y0=jac2+mysd,y1=jac2-mysd,code=3, angle=90, length=0.1 , col="gray40")
dev.off()

@@

#rearrange KEGG, Hallmark, HPO, Wikipath, miR, TFT, Reactome, Cellmarkers, GO
unsl <- unsl[,c(10,9, 6,5, 8,7, 18,17, 12,11, 16,15, 14,13, 2,1, 4,3)]

par(mar = c(5.1, 8.1, 4.1, 2.1))
library(beeswarm)
boxplot(unsl,horizontal=TRUE,las=1,col=c("pink","lightblue"),cex=0,xlab="no. significant sets FDR<0.05")
beeswarm(unsl,horizontal=TRUE,add=TRUE,col=c("red","blue"),pch=19,cex=0.6)

unsl1 <- unsl+1
boxplot(unsl1,horizontal=TRUE,log="x",las=1,col=c("pink","lightblue"),cex=0,xlab="no. significant sets FDR<0.05*")
beeswarm(unsl1,horizontal=TRUE,add=TRUE,col=c("red","blue"),pch=19,cex=0.6)

legend("bottomright", inset=.02, legend=c("corrected","original"),
  fill=c("lightblue","pink"),col=c("blue","red"), horiz=FALSE, cex=1.1)

png("fig1_ORAsigbox.png")
boxplot(unsl1,horizontal=TRUE,log="x",las=1,col=c("pink","lightblue"),cex=0,xlab="no. significant sets FDR<0.05*")
beeswarm(unsl1,horizontal=TRUE,add=TRUE,col=c("red","blue"),pch=19,cex=0.6)
legend("bottomright", inset=.02, legend=c("corrected","original"),
  fill=c("lightblue","pink"),col=c("blue","red"), horiz=FALSE, cex=1.1)
dev.off()

pdf("fig1_ORAsigbox.pdf",width=7, height=7)
par(mar = c(5.1, 8.1, 4.1, 2.1))
boxplot(unsl1,horizontal=TRUE,log="x",las=1,col=c("pink","lightblue"),cex=0,xlab="no. significant sets FDR<0.05*")
beeswarm(unsl1,horizontal=TRUE,add=TRUE,col=c("red","blue"),pch=19,cex=0.6)
legend("bottomright", inset=.02, legend=c("corrected","original"),
  fill=c("lightblue","pink"),col=c("blue","red"), horiz=FALSE, cex=1.1)
dev.off()

### 2024-09-09 analysis/analyse_rnaseq2.Rmd
adding error bars

# GO with error bars
plot(jac[,"GO"],type="b", pch=19,cex=1.5,lwd=2,
  xaxt = "n",ylim=c(0,1),ylab="Jaccard index",xlab="no. genes selected")
axis(1, at = seq(1, 5, by = 1),
     labels = ngenes)
mysd <- jsd[,"GO"]
arrows(x0=1:5, y0=(jac[,"GO"] - mysd) ,  x1=1:5, y1=(jac[,"GO"] + mysd), code=3, angle=90, length=0.1 )

mysd <- jsd[,"Reactome"]
points(jac[,"Reactome"],type="b",col="red",pch=19,cex=1.5,lwd=2)
arrows(x0=1:5, y0=(jac[,"Reactome"] - mysd) ,  x1=1:5, y1=(jac[,"Reactome"] + mysd), code=3, angle=90, length=0.1 ,col="red")

mysd <- jsd[,"Cellmarkers"]
points(jac[,"Cellmarkers"],type="b",col="blue",pch=19,cex=1.5,lwd=2)
arrows(x0=1:5, y0=(jac[,"Cellmarkers"] - mysd) ,  x1=1:5, y1=(jac[,"Cellmarkers"] + mysd), code=3, angle=90, length=0.1 , col="blue")

mysd <- jsd[,"TFT GTRD"]
points(jac[,"TFT GTRD"],type="b",col="gray50",pch=19,cex=1.5,lwd=2)
arrows(x0=1:5, y0=(jac[,"TFT GTRD"] - mysd) ,  x1=1:5, y1=(jac[,"TFT GTRD"] + mysd), code=3, angle=90, length=0.1 , col="gray50" )

plot(jac[,"GO"],type="b", pch=19,cex=1.5,lwd=2,
  xaxt = "n",ylim=c(0,1),ylab="Jaccard index",xlab="no. genes selected")
axis(1, at = seq(1, 5, by = 1),
     labels = ngenes)
points(jac[,"Reactome"],type="b",col="red",pch=19,cex=1.5,lwd=2)
points(jac[,"Cellmarkers"],type="b",col="blue",pch=19,cex=1.5,lwd=2)
points(jac[,"TFT GTRD"],type="b",col="gray50",pch=19,cex=1.5,lwd=2)

@@

>>>>

# with error bars 2x2 charts
ubtrim <- function(x) { if(is.na(x)) { x} else if(x > 1) { 1 } else { x } }
lbtrim <- function(x) { if(is.na(x)) { x} else if(x < 0) { 0 } else { x } }

par(mfrow=c(2,2))

# GO
plot(jac[,"GO"],type="b", pch=19,cex=1.5,lwd=2, xaxt = "n",ylim=c(0,1),ylab="Jaccard index",xlab="no. genes selected",main="GO")
axis(1, at = seq(1, 5, by = 1), labels = ngenes)
mysd <- jsd[,"GO"]
ub <- (jac[,"GO"] + mysd)
ub <- sapply(ub,ubtrim)
lb <-(jac[,"GO"] - mysd)
lb <- sapply(lb,lbtrim)
arrows(x0=1:5, y0=lb ,  x1=1:5, y1=ub, code=3, angle=90, length=0.1 )

# REACTOME
mysd <- jsd[,"Reactome"]
plot(jac[,"Reactome"],type="b", col="red", pch=19,cex=1.5,lwd=2, xaxt = "n",ylim=c(0,1),ylab="Jaccard index",xlab="no. genes selected",main="Reactome")
axis(1, at = seq(1, 5, by = 1), labels = ngenes)
ub <- (jac[,"Reactome"] + mysd)
ub <- sapply(ub,ubtrim)
lb <-(jac[,"Reactome"] - mysd)
lb <- sapply(lb,lbtrim)
arrows(x0=1:5, y0=lb ,  x1=1:5, y1=ub, code=3, angle=90, length=0.1 ,col="red")

# Cellmarkers
mysd <- jsd[,"Cellmarkers"]
plot(jac[,"Cellmarkers"],type="b", col="blue", pch=19,cex=1.5,lwd=2, xaxt = "n",ylim=c(0,1),ylab="Jaccard index",xlab="no. genes selected",main="Cellmarkers")
mysd <- jsd[,"GO"]
ub <- (jac[,"Cellmarkers"] + mysd)
ub <- sapply(ub,ubtrim)
lb <-(jac[,"Cellmarkers"] - mysd)
lb <- sapply(lb,lbtrim)
axis(1, at = seq(1, 5, by = 1), labels = ngenes)
arrows(x0=1:5, y0=lb,  x1=1:5, y1=ub, code=3, angle=90, length=0.1 , col="blue")

# TFT
mysd <- jsd[,"TFT GTRD"]
plot(jac[,"TFT GTRD"],type="b", col="gray50", pch=19,cex=1.5,lwd=2, xaxt = "n",ylim=c(0,1),ylab="Jaccard index",xlab="no. genes selected",main="TFT GTRD")
axis(1, at = seq(1, 5, by = 1), labels = ngenes)
ub <- (jac[,"TFT GTRD"] + mysd)
ub <- sapply(ub,ubtrim)
lb <-(jac[,"TFT GTRD"] - mysd)
lb <- sapply(lb,lbtrim)
arrows(x0=1:5, y0=lb,  x1=1:5, y1=ub, code=3, angle=90, length=0.1 , col="gray50" )

png("fig2_ORAszsd.png")
# GO
plot(jac[,"GO"],type="b", pch=19,cex=1.5,lwd=2, xaxt = "n",ylim=c(0,1),ylab="Jaccard index",xlab="no. genes selected",main="GO")
axis(1, at = seq(1, 5, by = 1), labels = ngenes)
mysd <- jsd[,"GO"]
ub <- (jac[,"GO"] + mysd)
ub <- sapply(ub,ubtrim)
lb <-(jac[,"GO"] - mysd)
lb <- sapply(lb,lbtrim)
arrows(x0=1:5, y0=lb ,  x1=1:5, y1=ub, code=3, angle=90, length=0.1 )

# REACTOME
mysd <- jsd[,"Reactome"]
plot(jac[,"Reactome"],type="b", col="red", pch=19,cex=1.5,lwd=2, xaxt = "n",ylim=c(0,1),ylab="Jaccard index",xlab="no. genes selected",main="Reactome")
axis(1, at = seq(1, 5, by = 1), labels = ngenes)
ub <- (jac[,"Reactome"] + mysd)
ub <- sapply(ub,ubtrim)
lb <-(jac[,"Reactome"] - mysd)
lb <- sapply(lb,lbtrim)
arrows(x0=1:5, y0=lb ,  x1=1:5, y1=ub, code=3, angle=90, length=0.1 ,col="red")

# Cellmarkers
mysd <- jsd[,"Cellmarkers"]
plot(jac[,"Cellmarkers"],type="b", col="blue", pch=19,cex=1.5,lwd=2, xaxt = "n",ylim=c(0,1),ylab="Jaccard index",xlab="no. genes selected",main="Cellmarkers")
mysd <- jsd[,"GO"]
ub <- (jac[,"Cellmarkers"] + mysd)
ub <- sapply(ub,ubtrim)
lb <-(jac[,"Cellmarkers"] - mysd)
lb <- sapply(lb,lbtrim)
axis(1, at = seq(1, 5, by = 1), labels = ngenes)
arrows(x0=1:5, y0=lb,  x1=1:5, y1=ub, code=3, angle=90, length=0.1 , col="blue")

# TFT
mysd <- jsd[,"TFT GTRD"]
plot(jac[,"TFT GTRD"],type="b", col="gray50", pch=19,cex=1.5,lwd=2, xaxt = "n",ylim=c(0,1),ylab="Jaccard index",xlab="no. genes selected",main="TFT GTRD")
axis(1, at = seq(1, 5, by = 1), labels = ngenes)
ub <- (jac[,"TFT GTRD"] + mysd)
ub <- sapply(ub,ubtrim)
lb <-(jac[,"TFT GTRD"] - mysd)
lb <- sapply(lb,lbtrim)
arrows(x0=1:5, y0=lb,  x1=1:5, y1=ub, code=3, angle=90, length=0.1 , col="gray50" )
dev.off()

pdf("fig2_ORAszsd.pdf",width=7, height=7)
par(mfrow=c(2,2))
# GO
plot(jac[,"GO"],type="b", pch=19,cex=1.5,lwd=2, xaxt = "n",ylim=c(0,1),ylab="Jaccard index",xlab="no. genes selected",main="GO")
axis(1, at = seq(1, 5, by = 1), labels = ngenes)
mysd <- jsd[,"GO"]
ub <- (jac[,"GO"] + mysd)
ub <- sapply(ub,ubtrim)
lb <-(jac[,"GO"] - mysd)
lb <- sapply(lb,lbtrim)
arrows(x0=1:5, y0=lb ,  x1=1:5, y1=ub, code=3, angle=90, length=0.1 )

# REACTOME
mysd <- jsd[,"Reactome"]
plot(jac[,"Reactome"],type="b", col="red", pch=19,cex=1.5,lwd=2, xaxt = "n",ylim=c(0,1),ylab="Jaccard index",xlab="no. genes selected",main="Reactome")
axis(1, at = seq(1, 5, by = 1), labels = ngenes)
ub <- (jac[,"Reactome"] + mysd)
ub <- sapply(ub,ubtrim)
lb <-(jac[,"Reactome"] - mysd)
lb <- sapply(lb,lbtrim)
arrows(x0=1:5, y0=lb ,  x1=1:5, y1=ub, code=3, angle=90, length=0.1 ,col="red")

# Cellmarkers
mysd <- jsd[,"Cellmarkers"]
plot(jac[,"Cellmarkers"],type="b", col="blue", pch=19,cex=1.5,lwd=2, xaxt = "n",ylim=c(0,1),ylab="Jaccard index",xlab="no. genes selected",main="Cellmarkers")
mysd <- jsd[,"GO"]
ub <- (jac[,"Cellmarkers"] + mysd)
ub <- sapply(ub,ubtrim)
lb <-(jac[,"Cellmarkers"] - mysd)
lb <- sapply(lb,lbtrim)
axis(1, at = seq(1, 5, by = 1), labels = ngenes)
arrows(x0=1:5, y0=lb,  x1=1:5, y1=ub, code=3, angle=90, length=0.1 , col="blue")

# TFT
mysd <- jsd[,"TFT GTRD"]
plot(jac[,"TFT GTRD"],type="b", col="gray50", pch=19,cex=1.5,lwd=2, xaxt = "n",ylim=c(0,1),ylab="Jaccard index",xlab="no. genes selected",main="TFT GTRD")
axis(1, at = seq(1, 5, by = 1), labels = ngenes)
ub <- (jac[,"TFT GTRD"] + mysd)
ub <- sapply(ub,ubtrim)
lb <-(jac[,"TFT GTRD"] - mysd)
lb <- sapply(lb,lbtrim)
arrows(x0=1:5, y0=lb,  x1=1:5, y1=ub, code=3, angle=90, length=0.1 , col="gray50" )
dev.off()

par(mfrow=c(1,1))

# original
plot(jac[,"GO"],type="b", pch=19,cex=1.5,lwd=2,
  xaxt = "n",ylim=c(0,1),ylab="Jaccard index",xlab="no. genes selected")
axis(1, at = seq(1, 5, by = 1),
     labels = ngenes)
points(jac[,"Reactome"],type="b",col="red",pch=19,cex=1.5,lwd=2)
points(jac[,"Cellmarkers"],type="b",col="blue",pch=19,cex=1.5,lwd=2)
points(jac[,"TFT GTRD"],type="b",col="gray50",pch=19,cex=1.5,lwd=2)

@@

Boxplot.

```{r,boxplot}

colnames(res3a) <- paste(colnames(res3a),"orig")
colnames(res3b) <- paste(colnames(res3b),"corr")
res3d <- cbind(res3a,res3b)

res3d <- as.data.frame(res3d[,c(7,16, 1,10, 9,18, 3,12, 4,13, 5,14, 8,17, 2,11, 6,15)])

par(mar = c(5.1, 8.1, 4.1, 2.1))
library(beeswarm)
boxplot(res3d,horizontal=TRUE,las=1,col=c("pink","lightblue"),cex=0,xlab="no. significant sets FDR<0.05")
beeswarm(res3d,horizontal=TRUE,add=TRUE,col=c("red","blue"),pch=19,cex=0.6)

res3d1 <- res3d + 1
boxplot(res3d1,log="x",horizontal=TRUE,las=1,col=c("pink","lightblue"),cex=0,xlab="no. significant sets FDR<0.05*")
beeswarm(res3d1,horizontal=TRUE,add=TRUE,col=c("red","blue"),pch=19,cex=0.6)

legend("bottomright", inset=.02, legend=c("corrected","original"),
  fill=c("lightblue","pink"),col=c("blue","red"), horiz=FALSE, cex=1.1)

png("fig2_ORAfdrbox.png")
par(mar=c(c(5.1, 7.1, 2.1, 2.1) ))

boxplot(res3d1,log="x",horizontal=TRUE,las=1,col=c("pink","lightblue"),cex=0,xlab="no. significant sets FDR<0.05*")
beeswarm(res3d1,horizontal=TRUE,add=TRUE,col=c("red","blue"),pch=19,cex=0.6)

legend("bottomright", inset=.02, legend=c("corrected","original"),
  fill=c("lightblue","pink"),col=c("blue","red"), horiz=FALSE, cex=1.1)

dev.off()

pdf("fig2_ORAfdrbox.pdf",width=7,height=7)
par(mar=c(c(5.1, 7.1, 2.1, 2.1) ))

boxplot(res3d1,log="x",horizontal=TRUE,las=1,col=c("pink","lightblue"),cex=0,xlab="no. significant sets FDR<0.05*")
beeswarm(res3d1,horizontal=TRUE,add=TRUE,col=c("red","blue"),pch=19,cex=0.6)

legend("bottomright", inset=.02, legend=c("corrected","original"),
  fill=c("lightblue","pink"),col=c("blue","red"), horiz=FALSE, cex=1.1)

dev.off()
